OLS with Interacting Variables

Population Density * Is Anti Wind Opinion

$$ \text{status}\ = \beta_0 + \beta_1 \text{pop_den}\_i + \beta_2 \text{med_inc} + \beta_3 \text{pop_den * med_inc} + \varepsilon_i $$

When wind turbines are operational, and all other variables are held constant at 0, the $\beta_0$ represents the estimated likelihood that status of a wind infrastructure plant is operating is ~ 4.46. $\beta_1$ tells us there is an estimated decrease of 2.79 in probability that status would be operating for the average unit increase in is_anti_wind opinion holders. For every average unit increase in pop_den, $\beta_2$ predicts 1.09e-04 will increase the operating status. The dramatic change in pop_den estimated influence on status is likely due to an interaction relationship with med_inc, which intuitively makes sense.

What is interaction telling us?

$\beta_3$ predicts 5.91e-05 will decrease the operating status, for an average unit increase in pop_den * is_anti_wind.

Our summary table describes a null deviance 362.60.(1178 df) and residual deviance 292.12 (1175 df). Additionally, it provides a p-value of 6.52e-14 for $\beta_1$, indicating that the hypothesis here is rejected because it is drastically smaller than the significant value of 0.05. The p-value of $\beta_2$, 0.837 is acceptable because it is within range of the significant value. The p-value of $\beta_3$, 0.931 is acceptable because it is within range of the significant value.


```{r}
# Regression with interacting beta
status4 <- glm(status ~ pop_den + med_inc + pop_den * med_inc, 
                       wind_us,
                       family = 'binomial')

summary(status4)
```

##### Population Density \* Median Income

FIX THE WORDS HERE

##### $$ \text{status}\ = \beta_0 + \beta_1 \text{pop_den}\_i + \beta_2 \text{median_age} + \beta_3 \text{pop_den * median_age} + \varepsilon_i $$

When wind turbines are operational, and all other variables are held constant at 0, the $\beta_0$ represents the estimated likelihood that `status` of a wind infrastructure plant is `operating` is \~ `4.23`.

$\beta_1$ tells us there is an estimated decrease of `2.81` in probability that status would be `operating` for the average unit increase in `is_anti_wind` opinion holders.

For every average unit increase in `pop_den`, $\beta_2$ predicts `1.65e-04` will increase the `operating status`, indicating `pop_den` is an extremely marginal influence, that may potentially be interacting with `is_anti_wind` outlook.

$\beta_3$ predicts `1.79e-05` will decrease the `operating status`, for an average unit increase in `med_inc`.

$\beta_4$ estimates an increase of `2.73e-02` in the `operating status`, for each average unit increase in `med_age`.

Our summary table describes a null deviance `362.60`.(`1178 df`) and residual deviance `289.02` (`1174 df`).

Additionally, it provides a p-value of `9.17e-15` for $\beta_1$, indicating that the hypothesis here is rejected because it is drastically smaller than the `significant value of 0.05`.

The p-value of $\beta_2$, `0.553` is acceptable because it is within range of the significant value.

The p-value of $\beta_3$, `0.096` is acceptable because it is within range of the significant value.

The p-value of $\beta_4$, \~`0.370` is acceptable because it is within range of the significant value.

```{r}
# Regression with interacting beta
status5 <- glm(status ~ pop_den + median_age + pop_den * median_age, 
                       wind_us,
                       family = 'binomial')

summary(status5)
```
#-----------------------

Assessing Omitted Variables Bias in OLS

OLS with 2 Variables

$$ \text{status}\ = \beta_0 + \beta_1 \text{pop_den}\_i + \beta_2 \text{is_anti_wind} + \varepsilon_i $$ When wind turbines are operational, and all other variables are held constant at 0, the $\beta_0$ represents the estimated likelihood that status of a wind infrastructure plant is operating is ~ 4.47. $\beta_1$ tells us there is an estimated decrease of 2.79 in probability that status would be operating for the average unit increase in is_anti_wind opinion holders. For every average unit increase in pop_den, $\beta_2$ predicts 7.6e-05 will increase the operating status, indicating pop_den is an extremely marginal influence, that may potentially be interacting with is_anti_wind outlook.

Our summary table describes a null deviance 362.60.(1178 df) and residual deviance 292.13 (1176 df). Additionally, it provides a p-value of 9.5e-15 for $\beta_1$, indicating that the hypothesis here is rejected because it is drastically smaller than the significant value of 0.05. The p-value of $\beta_2$, 0.819 is acceptable because it is within range of the significant value.

{r}
# Initial regression 2 betas for null
status2 <- glm(status ~ pop_den +  med_inc,
                       wind_us,
                       family = 'binomial')

summary(status2)

OLS with 3 Variables

$$ \text{status}\ = \beta_0 + \beta_1 \text{pop_den}\_i + + \beta_2 \text{med_inc} + \beta_3 \text{median_age} + \varepsilon_i $$ When wind turbines are operational, and all other variables are held constant at 0, the $\beta_0$ represents the estimated likelihood that status of a wind infrastructure plant is operating is ~ 5.35. $\beta_1$ tells us there is an estimated decrease of 2.80 in probability that status would be operating for the average unit increase in is_anti_wind opinion holders. For every average unit increase in pop_den, $\beta_2$ predicts 1.65e-04 will increase the operating status. The dramatic change in pop_den estimated influence on status is likely due to an interaction relationship with med_inc, which intuitively makes sense. $\beta_3$ predicts 1.74e-04 will decrease the operating status, for an average unit increase in med_inc.

Our summary table describes a null deviance 362.60.(1178 df) and residual deviance 289.8 (1175 df). Additionally, it provides a p-value of 1.03e-14 for $\beta_1$, indicating that the hypothesis here is rejected because it is drastically smaller than the significant value of 0.05. The p-value of $\beta_2$, 0.668 is acceptable because it is within range of the significant value. The p-value of $\beta_3$, 0.103 is acceptable because it is within range of the significant value.

When wind turbines are operational, and all other variables are held constant at 0, the $\beta_0$ represents the estimated likelihood that status of a wind infrastructure plant is operating is ~ 4.23. $\beta_1$ tells us there is an estimated decrease of 2.81 in probability that status would be operating for the average unit increase in is_anti_wind opinion holders. For every average unit increase in pop_den, $\beta_2$ predicts 1.65e-04 will increase the operating status, indicating pop_den is an extremely marginal influence, that may potentially be interacting with is_anti_wind outlook. $\beta_3$ predicts 1.79e-05 will decrease the operating status, for an average unit increase in med_inc. $\beta_4$ estimates an increase of 2.73e-02 in the operating status, for each average unit increase in med_age.

Our summary table describes a null deviance 362.60.(1178 df) and residual deviance 289.02 (1174 df). Additionally, it provides a p-value of 9.17e-15 for $\beta_1$, indicating that the hypothesis here is rejected because it is drastically smaller than the significant value of 0.05. The p-value of $\beta_2$, 0.553 is acceptable because it is within range of the significant value. The p-value of $\beta_3$, 0.096 is acceptable because it is within range of the significant value. The p-value of $\beta_4$, ~0.370 is acceptable because it is within range of the significant value.

{r}
# Regression with 3 beta
status3 <- glm(status ~ pop_den + med_inc + median_age, 
                       wind_us,
                       family = 'binomial')

summary(status3)

Finding P

solve for p using R.2

$logit(p)=log(p/1−p)=β0+β1x+ε$

toSolve <- function(p) {(1-p)*exp(1.56-0.05847*64) - p}

uniroot(toSolve, interval = c(0, 1))$root


Constructing Confidence Interval

Best Fit OLS Model 6

Model 6 is restricted to 1173 degrees of freedom, as output by our earlier OLS exploration. We will apply that value when constructing our 95% Confidence Intervals.

{r}
# Constructing a 95% confidence interval

# critical value for 2.5% quantile
c_val = qt(0.025, df = 1173, lower.tail = FALSE) %>% 
  print()

# confidence interval bounds
# since we saved our ols models as objects earlier, we can call them here

# using coef & SE from above to make 95% CI 

ci_lower <- round(status6$coefficients[2,'Estimate'] - c_val*status6$coefficients[2,'Std. Error'], 3) %>% 
  print()

ci_upper <- round(status6$coefficients[2,'Estimate'] + c_val*status6$coefficients[2,'Std. Error'], 3) %>% 
  print()

print("The first print statment refers to intercept confidence interval, and the second print statement refers to the slope CI")
print(paste0("95% probability that [", ci_lower_murder, " - ", ci_upper_murder, "] contains the difference in murder rates across the mean number of days that regions experienced frost vs no frost experience"))


#ci_lower <- round(point_est - c_val*SE, 2) %>% 
 # print()
#ci_upper <- round(point_est + c_val*SE, 2) %>% 
 # print()

# Display output
print(paste0("95% probability that interval [",ci_lower," ", ci_upper, "] contains the true difference in mean frost days between the North Central and South region."))


{r}



Logit

Null Hypothesis

After considering the glm

The null states, that the wind turbine active status is unimpacted by the is_anti_wind rating.

Two hypothesis presented are

$H_0: \text{is_anti_wind} - \text{not_anti_wind} = 0$

$H_A: \text{is_anti_wind} - \text{not_anti_wind} ≠ 0$

#---------------idk

{r}
ggplot(data= wind_us, aes(x = status, y = pop_den)) + geom_jitter(alpha=0.1, size=3) + geom_smooth(method='lm', formula= y~x, color="lightcoral", se=F, size=1.5) + theme_bw() + geom_hline(yintercept=0, color="seagreen")


Interpretting OVB
$\text{Null Deviance} = \text{2(LL(Saturated Model)} - \text{LL(Null Model)) on df} = \text{df_Sat - df_Null}$

Null deviance: tells us how well the response variable can be predicted by a model with only an intercept term.

$\text {Residual Deviance} = \text{2(LL(Saturated Model)} - \text{LL(Proposed Model))} \text{df} = \text {df_Sat} - \text{df_Proposed}$

Residual Deviance: tells us how well the response variable can be predicted by a model with p predictor variables.

$\text {The Saturated Model is a model that assumes each data point has its own parameters (which means you have n parameters to estimate.)}$

$\text {The Null Model assumes the exact "opposite", in that is assumes one parameter for all of the data points, which means you only estimate 1 parameter.}$

$\text{The Proposed Model assumes you can explain your data points with p parameters + an intercept term, so you have p+1 parameters.}$

When the Null Deviance is really small, it means that the Null Model explains the data pretty well. The same goes for Residual Deviance[1]. To best interpret our models, we will compare the Residual deviance.

Comparing Models using anova

Comparing our 6 residual deviance outputs, models 6 and 4 produced the smallest residual deviance at 288.84 and 289.02 respectively. To determine which is the best fit, we will do an anova assessment.

{r}
# this takes into consideration the predicted p value using both the null and residual deviation of each model and determines which more accurately describes the data
model_comparison <- anova(status6, status4, test = "Chisq") 

model_comparison

Our best fit OLS model is our interaction model 6: $$ \text{status}\ = \beta_0 + \beta_1 \text{is_anti_wind}\_i + + \beta_2 \text{pop_den} + \beta_3 \text{med_inc} + \beta_4 \text{med_age} + \varepsilon_i $$
