OLS with Interacting Variables

Population Density * Is Anti Wind Opinion

$$ \text{status}\ = \beta_0 + \beta_1 \text{pop_den}\_i + \beta_2 \text{med_inc} + \beta_3 \text{pop_den * med_inc} + \varepsilon_i $$

When wind turbines are operational, and all other variables are held constant at 0, the $\beta_0$ represents the estimated likelihood that status of a wind infrastructure plant is operating is ~ 4.46. $\beta_1$ tells us there is an estimated decrease of 2.79 in probability that status would be operating for the average unit increase in is_anti_wind opinion holders. For every average unit increase in pop_den, $\beta_2$ predicts 1.09e-04 will increase the operating status. The dramatic change in pop_den estimated influence on status is likely due to an interaction relationship with med_inc, which intuitively makes sense.

What is interaction telling us?

$\beta_3$ predicts 5.91e-05 will decrease the operating status, for an average unit increase in pop_den * is_anti_wind.

Our summary table describes a null deviance 362.60.(1178 df) and residual deviance 292.12 (1175 df). Additionally, it provides a p-value of 6.52e-14 for $\beta_1$, indicating that the hypothesis here is rejected because it is drastically smaller than the significant value of 0.05. The p-value of $\beta_2$, 0.837 is acceptable because it is within range of the significant value. The p-value of $\beta_3$, 0.931 is acceptable because it is within range of the significant value.


```{r}
# Regression with interacting beta
status4 <- glm(status ~ pop_den + med_inc + pop_den * med_inc, 
                       wind_us,
                       family = 'binomial')

summary(status4)
```

##### Population Density \* Median Income

FIX THE WORDS HERE

##### $$ \text{status}\ = \beta_0 + \beta_1 \text{pop_den}\_i + \beta_2 \text{median_age} + \beta_3 \text{pop_den * median_age} + \varepsilon_i $$

When wind turbines are operational, and all other variables are held constant at 0, the $\beta_0$ represents the estimated likelihood that `status` of a wind infrastructure plant is `operating` is \~ `4.23`.

$\beta_1$ tells us there is an estimated decrease of `2.81` in probability that status would be `operating` for the average unit increase in `is_anti_wind` opinion holders.

For every average unit increase in `pop_den`, $\beta_2$ predicts `1.65e-04` will increase the `operating status`, indicating `pop_den` is an extremely marginal influence, that may potentially be interacting with `is_anti_wind` outlook.

$\beta_3$ predicts `1.79e-05` will decrease the `operating status`, for an average unit increase in `med_inc`.

$\beta_4$ estimates an increase of `2.73e-02` in the `operating status`, for each average unit increase in `med_age`.

Our summary table describes a null deviance `362.60`.(`1178 df`) and residual deviance `289.02` (`1174 df`).

Additionally, it provides a p-value of `9.17e-15` for $\beta_1$, indicating that the hypothesis here is rejected because it is drastically smaller than the `significant value of 0.05`.

The p-value of $\beta_2$, `0.553` is acceptable because it is within range of the significant value.

The p-value of $\beta_3$, `0.096` is acceptable because it is within range of the significant value.

The p-value of $\beta_4$, \~`0.370` is acceptable because it is within range of the significant value.

```{r}
# Regression with interacting beta
status5 <- glm(status ~ pop_den + median_age + pop_den * median_age, 
                       wind_us,
                       family = 'binomial')

summary(status5)
```
